\pkg{DNest4}: Diffusive Nested Sampling in \proglang{C++} and \proglang{Python}

### 2016-05-26 paper/ms.tex
Talking about proposals

This generates $m$, $b$, and $\sigma$ from their joint prior. In this case,
the priors are all independent, so this reduces to generating the parameters
each from their own prior distribution. For convenience, the \code{RNG} object
has \code{rand()} and \code{randn()} member functions to generate from
a Uniform$(0,1)$ and Normal$(0,1)$ distribution respectively.

After generating the parameters from the prior, another member function
\code{calculate\_mu()} is called, which computes the model prediction
$mx_i + b$ for each data point.

@@

This first chooses a random integer from $[0, 3)$ using the
\code{rand\_int(int)} member function of the \code{DNest4::RNG} class.
This determines which of the three parameters
$(m, b, \sigma)$ is modified.

\subsection{Proposals for single parameters}
The proposal for $m$ involves adding a perturbation to the current value
using the line \code{m += 1E3*rng.randh();}.

A challenge using MCMC for standard Nested Sampling is that the target
distribution is not static --- it gets compressed over time. Similarly, in
the first stage of DNS the target distribution gets compressed (as levels are
added), and in the second stage the target distribution is a mixture of
distributions that have been compressed to varying degrees.

@@

\begin{enumerate}
\item Generate $c \sim$ Cauchy$(0, 1)$
\item Generate $n \sim$ Normal$(0, 1)$
\item Set $x := 10^{1.5 - |c|}n$
\end{enumerate}
This generates values with a maximum scale of tens-hundreds, down to
the smallest positive value in double precision, by virtue
of the Cauchy distribution's heavy tails.

For convenience, the \code{RNG} class contains a member function \code{randh()}
to generate values from this distribution.

The factor of \code{1E3} is
included to account for the prior width.

>>>>

This generates $m$, $b$, and $\sigma$ from their joint prior. In this case,
the priors are all independent, so this reduces to generating the parameters
each from their own prior distribution. For convenience,
\pkg{DNest4} provides an \code{RNG} class to represent random number
generators. The \code{RNG} class is just a convenience wrapper for the
random number generators built into \proglang{C++11}.
As you might expect, there are
\code{rand()} and \code{randn()} member functions to generate
double precision values
from a Uniform$(0,1)$ and Normal$(0,1)$ distribution respectively.

@@

This first chooses a random integer from $[0, 3)$ using the
\code{rand\_int(int)} member function of the \code{DNest4::RNG} class.
This determines which of the three parameters
$(m, b, \sigma)$ is modified. In this example, there are no proposals
that modify more than one of the parameters at a time, and all proposals
are ``random walk'' proposals that add a perturbation
(drawn from a symmetric distribution) to the current value.

\subsection{Proposals for single parameters}
The proposal for $m$ involves adding a perturbation to the current
value using the line

\begin{CodeChunk}
\begin{CodeInput}
m += 1E3*rng.randh();
\end{CodeInput}
\end{CodeChunk}

A challenge using MCMC for standard Nested Sampling is that the target
distribution is not static --- it gets compressed over time. Similarly, in
the first stage of DNS the target distribution gets compressed (as levels are
added), and in the second stage the target distribution is a mixture of
distributions that have been compressed to varying degrees.
This makes it difficult to tune step-sizes as you would when using
the standard Metropolis algorithm to sample the posterior distribution.

@@

\begin{enumerate}
\item Generate $c \sim$ Cauchy$(0, 1)$
\item Generate $n \sim$ Normal$(0, 1)$
\item Set $x := 10^{1.5 - |c|}n$
\end{enumerate}
This generates values with a maximum scale of tens--hundreds, down to
the smallest positive value in double precision, by virtue
of the Cauchy distribution's heavy tails.
For convenience, the \code{RNG} class contains a member function \code{randh()}
to generate values from this distribution.
The factor of \code{1E3} is included because it is a measure of the prior
width. Therefore, this proposal will attempt moves whose maximum order of
magnitude is a bit wider than the prior (since it would be very surprising
if any bigger moves were needed), and will propose moves a few orders of
magnitude smaller with moderate probability. We recommend using
this strategy (a measure of prior width, multiplied by \code{randh()}) as
a default proposal that works well in almost all problems.
